---
title: "ggDoubleHeat"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

The `ggDoubleHeat` package is a `ggplot2` extension that provides visualization for data from two different sources on a modified heat map. All functions from the package are named as `geom_heat_*()`. A regular heat map, which can be made by using `ggplot2::geom_tile()`, contains three dimensions (variables). `geom_heat_*()`, however, can express four dimensions of data on a single plot. 

A key component of `ggDoubleHeat` uses `ggnewscale`. If you are interested in learning it, check out its Github page (here is the [link](https://github.com/eliocamp/ggnewscale)).


## Installation

You can install the released version of `ggDoubleHeat` from the development version with:

``` r
# install.packages("devtools")
devtools::install_github("PursuitOfDataScience/ggDoubleHeat")
```

## Usage

Here in this section, two functions `geom_heat_grid()` and `geom_heat_circle()` are illustrated.

For demostration purposes, the built-in dataset `pitts_tg` to illustrate the basic usage of the package.

```{r}
library(ggDoubleHeat)
library(ggplot2)
library(ggnewscale)
pitts_tg
```

`pitts_tg` is a dataset that collects the 30-week period of Covid-related Google & Twitter incidence rate for 9 different categories from the Pittsburgh Metropolitan Statistical Area (MSA). For the complete information of the dataset, please type `?pitts_tg` on the console. 

`geom_heat_grid()`:

```{r}
ggplot(data = pitts_tg, aes(x = week, y = category)) +
  geom_heat_grid(outside = Google, inside = Twitter) +
  ggtitle("Pittsburgh Google & Twitter Incidence Rate (%) Comparison")
```

Now changing `geom_heat_grid()` to `geom_heat_circle()`,

```{r}
ggplot(data = pitts_tg, aes(x = week, y = category)) +
  geom_heat_circle(outside = Google, inside = Twitter) +
  ggtitle("Pittsburgh Google & Twitter Incidence Rate (%) Comparison")
  
```



